Robust Control of a Fluxonium Qubit

### 2020-11-16 paper/2background.tex
full build

%% S2
\section{QOC + AL-iLQR}
(QOC Problem Statement) Here we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schroedinger equation (TDSE)
\label{eq:tdse}
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
\end{align}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators and integrators of the Runge-Kutta type

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
the infidelity between the inital state evolved
to the final knot point ($\ket{\psi_{N}(u(t))}$)
and the target state ($\ket{\psi_{f}}$). In general
$J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional ($\nabla_{u} J(u)$).
They employ a first-order optimizer to minimize $J$ with respect to $u$.

(AL-iLQR Problem Statement) The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the quantum optimal control problem. The quantum optimal control
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. In the LQR formulation
the same functional is evaluated at each knot point
\begin{equation}
  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
\end{equation}
where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
at knot point $k$ and final state, $u_{k}$ are the controls,
and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
The state is propagated using a dynamics function
$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics. In the following
we refer to $\ket{\psi}$ as the state, $x$ as the augmented
state, and $u$ as the augmented controls.

The advantage of the LQR formulation
is that there exists a dynamic programming algorithm to compute the
optimal update to the augmented controls ($u_{k}$) which minimizes the functional
($J_{\textrm{iLQR}, k}$) for each knot point. This algorithm proceeds by deriving a
recurrence relation between knot points $k$ and $k + 1$ for the optimal
feedback law--known as the Ricatti recursion (see Appendix). The
iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
and applies the Ricatti recursion to all knot points on multiple
executions.

In order to incorporate constraints we employ
the augmented Lagrangian method. Constraints are contributions
to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
zero or negative when the constraint is satisfied. The AL-iLQR
method associates a penalty multiplier with the functional
that estimates the constraint's Lagrange multiplier.
The algorithm updates the penalty multiplier between
iLQR executions. In this scheme the functional takes the form
\begin{equation}
  \begin{aligned}
    J_{\textrm{AL-iLQR}} = \ &(\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
    &+ J_{\textrm{iLQR}}
  \end{aligned}
\end{equation}
where $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
with $\mu_{k}$ along the diagonal.
$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
\begin{align}
  \lambda_{k_{i}} &\gets \max(0, \lambda_{k_{i}} + \mu_{k_{i}} c_{k_{i}}(x_{k}^{*}, u_{k}^{*}))\\
  \mu_{k_{i}} &\gets \phi \mu_{k_{i}}
\end{align}
where $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
$i$ indicates the $i$-th constraint functional,
and $\phi$ is a hyperparameter. With this updated form of the cost
functional there still exists a recurrence relation to calculate the optimal control
updates, see \cite{howell2019altro}.

>>>>

\section{QOC + AL-iLQR}
In this section we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schr{\"o}dinger equation (TDSE)
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi} = H(u(t), t) \ket{\psi}
  \label{eq:tdse}
\end{equation}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system. Numerical quantum optimal control techniques make
the problem tractable by discretizing the problem into $N$
knot points (time steps). Typical explicit integration techniques for the TDSE include
exponential integrators \cite{auer2018magnus, berland2005solving, einkemmer2017performance}
and Runge-Kutta integrators \cite{jorgensen2011numerical}.

Quantum optimal control seeks the control
parameters that minimize a functional $J(u(t))$.
In the simplest case the functional is
the infidelity between the inital state evolved
to the final knot point and the target state
$J = 1 - {\lvert \braket{\psi_{f} \lvert \psi_{N}(u(t))} \rvert}^{2}$
In general $J$ is a linear combinaion of cost functions on the state, e.g.
forbidden-state occupation, as well as
cost functions on the controls, e.g. the norm of the control amplitudes
\cite{leung2017speedup}. Typical quantum optimal control
algorithms employ automatic differentiation
to compute first order information for the functional $\nabla_{u} J(u)$.
This information is used in a first-order optimizer to minimize $J$ with respect to $u$.
This scheme lends itself to projected gradient methods
which restrict the optimization to the constraint manifold
\cite{clarkson2010coresets, hauswirth2016projected}.

The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the QOC problem. The QOC
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. A general LQR cost
function takes the form
\begin{equation}
  J_{\textrm{iLQR}} = \ell_{N}({x}_{N})
  + \sum_{k = 0}^{N - 1} \ell_{k}({x}_{k}, u_{k})
\end{equation}
The subscript $k$ denotes the knot point, $x$ is the augmented state vector,
and $u$ is the augmented control vector. The augmented state is propagated
to the next knot point by the dynamics function $x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of QOC $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics \eqref{eq:tdse}. The
cost function $\ell_{k}$ includes contributions from the infidelity
$1 - {\lvert \braket{\psi_{f} \lvert \psi_{k}} \rvert}^{2}$, final state deviations
$\tilde{x}^{T}_{k} Q_{k} \tilde{x}_{k}$ where $\tilde{x}_{k} = x_{f} - x_{k}$,
and control regularization $u^{T}_{k} R_{k} u_{k}$.
Placing cost functions at each knot point smoothens the optimization
landscape. In QOC realizing the final state
is the most important objective so the cost function at the final knot point
carries a higher realtive weight $\ell_{N} \sim N \cdot \ell_{k}$.
The LQR formulation restricts the cost functions to operate on the augmented
state and controls at each knot point. The advantage of the LQR formulation
is that there exists a differentiable dynamic programming algorithm to compute the
optimal update to the augmented controls $u_{k}$ which minimizes the cost
function $\ell_{k}$ at each knot point. This algorithm proceeds by deriving a
recurrence relation between knot points $k$ and $k + 1$ for the optimal
feedback law--known as the Ricatti recursion (citations needed). The
iterative LQR (iLQR) algorithm computes $J_{\textrm{iLQR}}$
and applies the Ricatti recursion to all knot points on multiple
executions.

In order to incorporate constraints we employ
the augmented Lagrangian method. Constraints are contributions
to the functional of arbitrary form $c_{k}(x_{k}, u_{k})$ which are
zero or negative when the constraint is satisfied. The AL-iLQR
method associates a penalty multiplier with the constraint
that estimates the constraint's Lagrange multiplier.
The algorithm updates the penalty multiplier between
iLQR executions. In this scheme the functional takes the form
\begin{equation}
\begin{aligned}
    J_{\textrm{AL-iLQR}} &= J_{\textrm{iLQR}}\\
    &+ (\lambda_{N} + \frac{1}{2}I_{\mu_{N}} c_{N}(x_{N}))^{T} c_{N}(x_{N})\\
    &+ \sum^{N - 1}_{k = 0} (\lambda_{k} + \frac{1}{2}I_{\mu_{k}} c_{k}(x_{k}, u_{k}))^{T} c_{k}(x_{k}, u_{k})\\
\end{aligned}
\end{equation}
Here $\lambda_{k}$ is a Lagrange multiplier and $I_{\mu_{k}}$ is a penalty matrix
with $\mu_{k}$ along the diagonal.
$\lambda_{k}$ and $\mu_{k}$ are updated after each augmented Lagrangian iteration according to
\begin{align}
  \lambda^{i}_{k} &\gets \max(0, \lambda^{i}_{k} + \mu^{i}_{k} c^{i}_{k}(x_{k}^{*}, u_{k}^{*}))\\
  \mu^{i}_{k} &\gets \phi \mu^{i}_{k}
\end{align}
Here $x^{*}, u^{*}$ are the optimal augmented state and augmented controls from the iLQR execution,
$i$ indicates the $i$-th constraint functional,
and $\phi$ is a hyperparameter. An optimal control update, similar to the Ricatti
recursion, can be formed to calculate the optimal control updates for this functional,
see \cite{howell2019altro}.

### 2020-11-16 paper/3fluxonium.tex
full build

\section{QOC on the Fluxonium}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form
\label{eq:hamiltonian}
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \ref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$.
The equality and inequality constraints on $a$ are handled
with a bound constraint.
%% TODO: probably want to list equations for the constraints

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

%% TODO: talk about quadratic cost vs. infidelity

>>>>

\section{QOC on the Fluxonium}
In the following we study
the quantum optimal control problem on the fluxonium qubit.
The fluxonium qubit is a promising building block for superconducting
circuits, and the experimental constraints we encode in the optimization reflect
those of a realistic device. Furthermore, the accurate
two-level approximation of the system Hamiltonian makes it
efficient to perform quantum optimal control on
a classical computer. In the two-level
approximation the system Hamiltonian takes the form
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
  \label{eq:hamiltonian}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
systematic errors. We compare the gates we obtain with numerical
methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.

In addition to imposing optimization constraints that
reflect physical limitations of the apparatus, we impose
constraints that improve the experimental realization of the control pulse.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines,
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \eqref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.
The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$, the constraint functional takes the form
$c_{k} = (x_{k} - x_{f})^{T}(x_{k} - x_{f})$.
The equality and inequality constraints on $a$ are handled
with a bound constraint which takes the form
$c_{k} = (x_{k} - b)^{T}(x_{k} - b)$ if $\lvert x_{k} \rvert \ge b$
and $0$ otherwise.
In addition the norm of the quantum state
is constrained to ensure discretization error is not
exploited $c_{k} = {\lvert \braket{\psi_{k} \lvert \psi_{k}} \rvert}^{2} - 1$.

In addition to obeying experimental constraints
and achieving low simulated gate errors, we desire
to make our gates perform well in the presence of decoherence.
Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively.
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.
Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
by tuning the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

### 2020-11-16 paper/5static.tex
full build

%% TODO: standardize use of "paramter" vs "system parameter" language
\section{Robustness to Static Parameter Deviations \label{sec:static}}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
We consider three robust control techniques.
The first is the sampling method, which has
been proposed previously in the context of QOC
\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
sampling method, multiple states are evolved under distinct deviant dynamics
to capture the effect of parameter deviations. We also
study the unscented sampling method, which uses the unscented
transformation to accurately propagate a distribution
representing the uncertainty in an evolving state
due to a parameter deviation.
The unscented transformation was designed for nonlinear Kalman
filtering and is frequently utilized in robust control
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
Finally, we propose the derivative method. Derivative
information encoding the sensitivity of the
state trajectory with respect to the deviant parameter
is used to penalize state trajectory deviations.

In the sampling method, sample states evolve under a
Hamiltonian where a parameter is replaced by
a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
The gate error of each sample state is 
penalized at each knot point
$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
Each initial state is used to initialize two
sample states $\ket{\psi^{\pm}}$.
The initial states are chosen so
that their outer products span the operators on the
Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
This is a standard perscription, though knowledge
of the problem can typically be used to reduced the number
of required samples. For this method the number of states in the augmented state vector
scales as $O(dn^{2})$ because each of $n^{2}$ initial
states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
and the size of the state vector. $d$ is the number of deviant parameters.

%% F3
\begin{figure*}[ht]
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}\hspace{0.025\textwidth}
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3b.png}
  \end{subfigure}
  \caption{
    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
    sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates shown
    are the solutions at the analytic gate time.
    (b) Simulation of stochastic 1/$f$ flux noise for
    successive gate applications. The reported gate error is the cumulative
    gate error after each gate application.
  }
  \label{fig:stochastic}
\end{figure*}

@@

The derivative method draws on the intuition that
the sensitivity of the state evolution to a parameter
$\lambda$ is encoded in the $l$\textsuperscript{th}-order
derivative of the state with respect to that parameter
$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
derivative method minimizes the norm of the first $m$
state derivatives with a quadratic cost at each knot point
${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
The state derivatives could be obtained with backwards mode differentiation.
Naive automatic differentiation would compute
the state derivative at all $1, \dots, k - 1$ knot points
to obtain the state derivative at knot point $k$.
For a single state derivative and $N$ knot points this
requires $O(N^2)$ matrix multiplications.
Instead, we forward propagate the state derivatives in the
augmented state vector under coupled dynamics, resulting in
$O(N)$ matrix multiplications. For example, the dynamics
for the 1\textsuperscript{st}-order derivative method are
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
  H \ket{\partial_{\lambda} \psi} +
  (\partial_{\lambda} H) \ket{\psi}
\end{align}
Exponential integrators that account for the non-linear
term may be used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}.
For this method $m$ state derivatives are associated with each
initial state from the operator basis.
So, the number of states in the augmented state vector scales as
$O(dmn^{2})$.

To demonstrate the applicability of these techniques
to mitigate system parameter deviations,
we consider the task of achieving a single $Z/2$
gate subject to a constant qubit frequency detuning
$f_{q} \gets f_{q} + \delta f_{q}$.
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling methods accordingly. For each method we compute the gate error for
one simulated gate application subject to the deviant dynamics given by the
stated qubit frequency detuning $\delta f_{q}$.

@@

%% TODO: how do I tie this in?
Multiple analytic methods have also been presented to construct gates
that mitigate the error arising from parameter deviations
including composite pulse sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, and
geometric phase considerations
\cite{xu2020nonadiabatic, han2020experimental}.
Composite pulse sequences are derived by computing
the erroneous rotation arising from the static parameter deviation
with a finite order magnus expansion.
Pulses are then suitably composed to eliminate the
error. In principle the error may be eliminated
to arbitrary order with sufficiently many pulses.
It is difficult to choose an appropriate composite pulse
for comparison on the problem we study here. We
propose comparing numerical methods to existing analytic
techniques for future work.

>>>>

\section{Robustness to Static Parameter Deviations \label{sec:static}}
We have formulated the quantum optimal control
problem as an open loop optimization problem, i.e.
feedback from the experiment is not incorporated in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter deviations. As an example
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency $\tilde{f_{q}} = f_{q} \pm \sigma_{f_{q}}$.
We consider three robust control techniques.
The first is the sampling method, which has
been proposed previously in the context of QOC
\cite{rembold2020introduction, reinhold2019controlling, carvalho2020error}. In the
sampling method, multiple states are evolved under distinct deviant dynamics
to capture the effect of parameter deviations. We also
study the unscented sampling method, which uses the unscented
transformation to accurately propagate a distribution
representing the uncertainty in an evolving state
due to a parameter deviation.
The unscented transformation was designed for nonlinear Kalman
filtering and is frequently utilized in robust control
\cite{howell2020direct, julier2004unscented, lee2013sigma, manchester2016derivative}.
Finally, we propose the derivative method. Derivative
information encoding the sensitivity of the
state trajectory with respect to the deviant parameter
is used to penalize state trajectory deviations.

In the sampling method, sample states evolve under a
Hamiltonian where a parameter is replaced by
a deviant value. We propagate the additional states $\ket{\psi^{\pm}}$ with
deviant values $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
The gate error of each sample state is 
penalized at each knot point
$1 - {\lvert \braket{\psi^{\pm}_{k} | \psi_{f}} \rvert}^{2}$.
Each initial state is used to initialize two
sample states $\ket{\psi^{\pm}}$.
The initial states are chosen so
that their outer products span the operators on the
Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}.
For this method the number of states in the augmented state vector
scales as $O(dn^{2})$ because each of $n^{2}$ initial
states is represented by $2d$ samples. $n$ is both the dimension of the Hilbert space
and the size of the state vector. $d$ is the number of deviant parameters.

%% F3
\begin{figure*}[ht]
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}\hspace{0.025\textwidth}
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f3b.png}
  \end{subfigure}
  \caption{
    (a) $X/2$ gates robust to flux offsets constructed with the analytic,
    sampling, unscented sampling, and the 1\textsuperscript{st}-
    and 2\textsuperscript{nd}-order derivative methods. The gates shown
    are the solutions at the analytic gate time.
    (b) Simulation of stochastic 1/$f$ flux noise for
    successive gate applications. The cumulative
    gate error is computed after each gate application.
  }
  \label{fig:stochastic}
\end{figure*}

@@

The derivative method draws on the intuition that
the sensitivity of the state evolution to a parameter
$\lambda$ is encoded in the $l$\textsuperscript{th}-order
derivative of the state with respect to that parameter
$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
derivative method minimizes the norm of the first $m$
state derivatives with a quadratic cost at each knot point
${\lvert \braket{\partial^{l}_{\lambda} \psi | \partial^{l}_{\lambda} \psi}
  \rvert}^{2}$, $l \in \{1, \dots, m\}$.
The state derivatives could be obtained with backwards mode differentiation.
Naive automatic differentiation would compute
the state derivative at all $1, \dots, k - 1$ knot points
to obtain the state derivative at knot point $k$.
For a single state derivative and $N$ knot points this
requires $O(N^2)$ matrix multiplications.
Instead, we forward propagate the state derivatives in the
augmented state vector under coupled dynamics, resulting in
$O(N)$ matrix multiplications. For example, the dynamics
for the 1\textsuperscript{st}-order derivative method are
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
  H \ket{\partial_{\lambda} \psi} +
  (\partial_{\lambda} H) \ket{\psi}
\end{align}
Exponential integrators that account for the non-linear
term may be used to efficiently integrate the coupled dynamics,
see Appendix \ref{appendix:derivative}.
For this method $m$ state derivatives are associated with each
initial state from the operator basis.
So, the number of states in the augmented state vector scales as
$O(dmn^{2})$. Additionally, we note that the number of initial
states and samples we have employed for each of the three
methods follows standard perscriptions. However, knowledge
of the problem can typically be used to reduced the number
of required samples.

To demonstrate the applicability of these techniques
to mitigate parameter deviations,
we consider the task of achieving a single $Z/2$
gate subject to a constant qubit frequency detuning
$f_{q} \gets f_{q} + \delta f_{q}$.
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling methods accordingly. For each method we compute the gate error for
one simulated gate application subject to the deviant dynamics given by the
stated qubit frequency detuning $\delta f_{q}$.

@@

Additionally, there are multiple analytic methods 
to mitigate errors due to parameter deviations
including composite pulse sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, and
geometric phase considerations
\cite{xu2020nonadiabatic, han2020experimental}.
Composite pulse sequences are derived by computing
the erroneous rotation arising from the parameter deviation
with a finite order magnus expansion.
Pulses are then suitably composed to eliminate the
error. In principle the error may be eliminated
to arbitrary order with sufficiently many pulses \cite{merill2014progress},
which mirrors the negative correlation between
gate error and gate duration we observe with the derivative method.
It is difficult to choose an appropriate composite pulse
for the problem studied here,
so we propose comparisons between composite pulses and numerical techniques
for future work.

### 2020-11-16 paper/6stochastic.tex
full build

We compare the numerical methods discussed in the previous section
on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
The flux noise is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}
\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its sampled deviations
follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
in the augmented state vector. In principle the basic sampling method could be modified
similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
methods require no modification from the static case.

We simulate successive applications of the gate constructed by each method and compute the cumulative gate error
after each application, see Figure \ref{fig:stochastic}. Both the analytic and numerical methods achieve single gate errors
sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
numerical methods perform similarly in the concatenated gate application comparision. They achieve a two
order of magnitude reduction in the total gate error over the analytic method after $200$
gate applications $\sim 11 \mu\textrm{s}$.
1/$f$ flux noise is a signficiant source of coherent errors in NISQ applications and
these numerical techniques offer effective avenues to mitigate them.

>>>>

We compare the analytic gate and those produced by
the numerical methods discussed in the previous section
on the task of realizing a $X/2$ gate subject to 1/$f$ flux noise.
The flux noise is generated by
filtering white noise sampled from a standard normal distribution with a finite
impulse response filter \cite{saspweb2011}
\footnote{\url{https://ccrma.stanford.edu/~jos/sasp/Example_Synthesis_1_F_Noise.html}}.
It is then scaled by the 
flux noise amplitude of our device $A_{\Phi} = 5.21 \mu \Phi_{0} \implies
\delta a \sim 2.5 \cdot 10^{-5} \textrm{GHz}$.
The unscented sampling method is modified so that its sampled deviations
follow a 1/$f$ distribution by carrying the state of a finite impulse response filter
in the augmented state vector. In principle the basic sampling method could be modified
similarly but we choose to sample statically at $\delta a$ for comparision. The derivative
methods require no modification from the static case.

We simulate successive applications of the gate constructed by each method
and compute the cumulative gate error
after each application, see Figure \ref{fig:stochastic}. Both the analytic
and numerical methods achieve single gate errors
sufficient for quantum error correction. Despite converging on qualitatively different solutions, the
numerical methods perform similarly in the concatenated gate application comparision. They achieve a two
order of magnitude cumulative gate error reduction over the analytic method after $200$
gate applications $\sim 11 \mu\textrm{s}$.
1/$f$ flux noise is a signficiant source of coherent errors in NISQ applications and
these numerical techniques offer effective avenues to mitigate them.
